### 2024-05-06 data/template.qmd
Initial sketch

# template

We code the studies in ascii format into files titled `authoryearxx.R` stored in `data/` folder.

## Fields

- Initial list, to be modified
- Issues that need to be resolved can be marked after hashtag (#).
- If there are multiple studies/samples in one article, another data can be created and `study_id` can be adjusted by adding a postfix (`_a`,`_b`,`_c`).

```{r}
#| eval: false
#| echo: true
study_id <- 'authorlastnameyeartwoletterfromthetitle'
emotion_type <- c('dimensional','discrete','music-specific','other')
emotion_list <- c('valence','arousal','happy','sad','anger','fear','tender','other')
emotion_locus <- c('perceived','induced','not specified')
stimulus_type <- c('classical','pop','genre','mixed')
stimulus_duration <- 20
stimulus_N <- 24
participant_N <- 35
participant_expertise <- c('non-musician','musician','not specified')
participant_origin <- c('Country','not-specified')
participant_sampling <- c('convenience','not-specified')
participant_task <- c('rating','forced choice','ranking','other')
feature_list <- c('attack rate','pitch height','mode')
feature_validation <- c('annotation','none') # this was done in another study
feature_source <- c('librosa','essentia','MIR toolbox','marsyas','custom','not specified')
model_type <- c('correlation','regression','commonality analysis','classification')
model_measure <- c('rsquared','cor','rmse','classification','other')
model_complexity_parameters <- c(numeric, 'not defined')
model_rate_emotion_list <- c(0.00,0.00)
model_validation <- c('none','cross-validation')
meta_comments <- 'template'
meta_encoder <- 'Initials of the name'
meta_date <- '3/5/2024'
```