### 2024-05-06 preregistration/preregistration.qmd

## Meta-analysis title: To what degree music emotion recognition

## Authors: TE + CA + MS ? (to be decided)

## Brief rationale

To assess what degree models are able to predict emotions in music, what models, what emotions, what genres, etc.

## Research objectives

1. How good are MER models?

2. What are the systematic biases of the models?

3. How are models dealing with genres and cultures?

4. How generalisible are the models

5. WEIRD

6. Theory orientations

FINER: feasible, interesting, novel, ethical, relevant [@browner2022de]

## Hypotheses 
We expect that the

## Eligibility criteria
English, peer-reviewed journal articles indexed in X and Y, published between 20xx-20xx

## Database sources

(PubMed, Web of Science, Scopus)

## Search strategy
Music emotion recognition

## Analysis plan

Fixed effects model
Tests of small study/publication bias