### 2024-05-06 studies/search_syntax.qmd

# Search Syntax

| Database       | Date of search | Results | Search syntax                                                                                                                                                                                                                                                                                                              |
| :--------------| :------------- | :------ | :----------------------------------------------------------------------------------------------------- |
| PsychINFO      | X May 20XX    | XXX     | (emotion OR mood* OR affect*) AND (music OR audio* OR recognition OR percei* OR recogn*) AND (music)   |
| Scopus         | X May 20XX    | XXX     |           |
| Web of Science | X May 20XX    | XXX     |           |