### 2024-05-13 studies/search_syntax.qmd
first pass of the study lists

# Search Syntax

| Database       | Date of search | Results | Search syntax                                                                                                                                                                                                                                                                                                              |
| :--------------| :------------- | :------ | :----------------------------------------------------------------------------------------------------- |
| PsychINFO      | X May 20XX    | XXX     | (emotion OR mood* OR affect*) AND (music OR audio* OR recognition OR percei* OR recogn*) AND (music)   |
| Scopus         | X May 20XX    | XXX     |           |
| Web of Science | X May 20XX    | XXX     |           |

## Scopus

TITLE-ABS-KEY ( valence OR arousal OR classi OR categor OR algorithm AND music AND emotion AND recognition ) AND PUBYEAR > 2013 AND PUBYEAR < 2025 AND ( LIMIT-TO ( DOCTYPE , "ar" ) )

227_ results
scopus_first_pass.bib

## Web of Science

(DT=(Article) AND PY=(2014-2025)) AND ALL=(music emotion recognition valence arousal)

142 hits

webofscience_142_first_pass.bib

## OpenAlex

https://openalex.org/works?page=1&filter=default.search%3Amusic%20emotion%20recognition%20valence%20arousal,type%3Atypes%2Farticle,publication_year%3A2014-2024,keywords.id%3Akeywords%2Femotion-recognition,keywords.id%3Akeywords%2Faffective-computing,language%3Alanguages%2Fen,open_access.any_repository_has_fulltext%3Atrue

278