### 2024-05-20 studies/pass3_comparison.qmd

---
title: "Pass 3 Comparison"
author: "Cameron Anderson"
format: html
---

## Reading in Biblibrary

## Load libraries

```{r}
# load dplyr
library(dplyr)
```

## Read data

```{r}
# read in bibtex library as data frame
bib_df_ca <- bib2df::bib2df('metaMER_lib_third_pass_clean.bib')

```

## Compare

```{r}
dim(bib_df_merged)
library(stringr)

bib_df_merged$NOTES_INDEX.CA<-NA
bib_df_merged$NOTES_INDEX.CA[str_detect(tolower(bib_df_merged$NOTES.CA),'include')]<-'include'
bib_df_merged$NOTES_INDEX.CA[str_detect(tolower(bib_df_merged$NOTES.CA),'exclude')]<-'exclude'
bib_df_merged$NOTES_INDEX.CA[str_detect(tolower(bib_df_merged$NOTES.CA),'unsure')]<-'unsure'

bib_df_merged$NOTES_INDEX.TE<-NA
bib_df_merged$NOTES_INDEX.TE[str_detect(tolower(bib_df_merged$NOTES.TE),'include')]<-'include'
bib_df_merged$NOTES_INDEX.TE[str_detect(tolower(bib_df_merged$NOTES.TE),'exclude')]<-'exclude'
bib_df_merged$NOTES_INDEX.TE[str_detect(tolower(bib_df_merged$NOTES.TE),'unsure')]<-'unsure'
sum(is.na(bib_df_merged$NOTES_INDEX.CA))
sum(is.na(bib_df_merged$NOTES_INDEX.TE))

t<-table(bib_df_merged$NOTES_INDEX.CA,bib_df_merged$NOTES_INDEX.TE)
t2<-round(t/sum(t),2)
sum(diag(t2))
```
## Report annotation reliability/agreement

## Resolve these discrepancies
```{r}
# resolve unsure ones
IND <- which(bib_df_merged$NOTES_INDEX.CA=='unsure')
resolved_index <- bib_df_merged$BIBTEXKEY[IND]
resolved <-c('exclude',
             'exclude',
             'exclude',
             'exclude',
             'exclude',
             'exclude',
             'exclude',
             'exclude',
             'exclude',
             'exclude',
             'exclude',
             'exclude')

# Resolving conflicting exclude/include annotations, part 1
IND <- which(bib_df_merged$NOTES_INDEX.CA=='exclude' & bib_df_merged$NOTES_INDEX.TE=='include')
#IND <- which(bib_df_merged$NOTES_INDEX.TE=='exclude' & bib_df_merged$NOTES_INDEX.CA=='include')
resolved_exclude_index <- bib_df_merged$BIBTEXKEY[IND]
resolved_exclude_index # 8
resolved_exclude <-c('exclude','exclude','exclude','include','exclude','exclude','exclude','exclude')

# Resolving conflicting exclude/include annotations, part 2
IND <- which(bib_df_merged$NOTES_INDEX.TE=='exclude' & bib_df_merged$NOTES_INDEX.CA=='include')
resolved_exclude2_index <- bib_df_merged$BIBTEXKEY[IND]
resolved_exclude2_index # 8
resolved_exclude2 <-c('exclude','exclude','exclude','exclude')

```

## Export bibtex for the annotation

Notes:

For Feature classification, use this one:

https://ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=9229494

Insert the template the annotation fields from googledoc to the bibtex file.