### 2024-05-21 studies/pass3_comparison.qmd

---
title: "Pass 3 Comparison"
author: "Cameron Anderson"
format: html
---

## Reading in Biblibrary

## Load libraries

```{r}
# load libraries
library(dplyr)
library(stringr)
library(knitr)
```

## Read data

```{r}
# read in bibtex library as data frame
bib_df_merged <- bib2df::bib2df('metaMER_library_third_pass_clean.bib')
```

## Compare

```{r}
# check dimensions are accurate
dim(bib_df_merged)

# distinguish notes with author initials
names(bib_df_merged)[names(bib_df_merged) == 'NOTES'] <- 'NOTES.CA'
names(bib_df_merged)[names(bib_df_merged) == 'NOTES.1'] <- 'NOTES.TE'

# extract decisions less comments
capture_group <- 'include|exclude|unsure'

# create new index to track entries
bib_df_merged$NOTES_INDEX.CA<-NA

# create new column tracking decisions
bib_df_merged$NOTES_INDEX.CA <- str_extract(tolower(bib_df_merged$NOTES.CA),
                                         capture_group)
bib_df_merged$NOTES_INDEX.TE <- str_extract(tolower(bib_df_merged$NOTES.TE),
                                     capture_group)

# check entries are consistent
sum(is.na(bib_df_merged$NOTES_INDEX.CA))
sum(is.na(bib_df_merged$NOTES_INDEX.TE))

```

## Report annotation reliability/agreement

```{r}
# compare raters' decisions with confusion matrix
t<-table(bib_df_merged$NOTES_INDEX.CA,
         bib_df_merged$NOTES_INDEX.TE)

# get agreement
t2<-round(t/sum(t),2)
ag_before <- sum(diag(t2))

# make table
knitr::kable(t,  
             caption = paste('Votes before discussion. \n
             Rows: CA votes; cols: TE votes, Agreement = ', ag_before)
             )
```

## CA: unsure

```{r}
# resolve unsure ones
IND <- which(bib_df_merged$NOTES_INDEX.CA=='unsure')
resolved_index <- bib_df_merged$BIBTEXKEY[IND] # 12
# CA updates ratings:
bib_df_merged$NOTES_INDEX.CA[bib_df_merged$BIBTEXKEY == 'deng2024an'] <- 'exclude'
bib_df_merged$NOTES_INDEX.CA[bib_df_merged$BIBTEXKEY == 'hao2022re'] <- 'exclude'
bib_df_merged$NOTES_INDEX.CA[bib_df_merged$BIBTEXKEY == 'he2022al'] <- 'exclude'
bib_df_merged$NOTES_INDEX.CA[bib_df_merged$BIBTEXKEY == 'huang2023th'] <- 'exclude'
bib_df_merged$NOTES_INDEX.CA[bib_df_merged$BIBTEXKEY == 'priscillajoy2023mu'] <- 'exclude'
bib_df_merged$NOTES_INDEX.CA[bib_df_merged$BIBTEXKEY == 'shen2024re'] <- 'exclude'
bib_df_merged$NOTES_INDEX.CA[bib_df_merged$BIBTEXKEY == 'wang2021mu'] <- 'exclude'
bib_df_merged$NOTES_INDEX.CA[bib_df_merged$BIBTEXKEY == 'wang2021re'] <- 'exclude'
bib_df_merged$NOTES_INDEX.CA[bib_df_merged$BIBTEXKEY == 'wang2022mua'] <- 'exclude'
bib_df_merged$NOTES_INDEX.CA[bib_df_merged$BIBTEXKEY == 'zhang2016rea'] <- 'exclude'

# both CA and TE update ratings:
bib_df_merged$NOTES_INDEX.CA[bib_df_merged$BIBTEXKEY == 'na2022mu'] <- 'exclude'
bib_df_merged$NOTES_INDEX.TE[bib_df_merged$BIBTEXKEY == 'na2022mu'] <- 'exclude'
bib_df_merged$NOTES_INDEX.CA[bib_df_merged$BIBTEXKEY == 'tian2023mu'] <- 'exclude'
bib_df_merged$NOTES_INDEX.TE[bib_df_merged$BIBTEXKEY == 'tian2023mu'] <- 'exclude'

# resolved <-c('deng2024an' = 'exclude',
#              'hao2022re' = 'exclude',
#              'he2022al' = 'exclude',
#              'huang2023th' = 'exclude',
#              'na2022mu' = 'exclude',
#              'priscillajoy2023mu' = 'exclude',
#              'shen2024re' = 'exclude',
#              'tian2023mu' = 'exclude',
#              'wang2021mu' = 'exclude',
#              'wang2021re' = 'exclude',
#              'wang2022mua' = 'exclude',
#              'zhang2016rea' = 'exclude')

```

## TE: unsure (in progress)

```{r}
# update TE decision
bib_df_merged$NOTES_INDEX.TE[bib_df_merged$BIBTEXKEY == 'tang2023ap'] <- 'exclude'
bib_df_merged$NOTES_INDEX.TE[bib_df_merged$BIBTEXKEY == 'xing2015em'] <- 'exclude'

# update TE and CA decisions
bib_df_merged$NOTES_INDEX.TE[bib_df_merged$BIBTEXKEY == 'wang2022mu'] <- 'exclude'
bib_df_merged$NOTES_INDEX.CA[bib_df_merged$BIBTEXKEY == 'wang2022mu'] <- 'exclude'

```
## Summary
|Study|CA|TE||Resolution/Decision|
|---|---|---|---|
|tang2023ap|uses image features|No relevant features, lack of stimulus detail|Exclude|
|xing2015em|includes classification task|No N of musical excerpts, missing information regarding data processing|Exclude|
|wang2022mu|only reports DET and equal error rate|Although they report correlation coefficients, emotions are not valence or arousal. Do report MSE of classification task|Exclude|

## Tabulate results after resolving unsures

```{r}
# compare raters' decisions with confusion matrix
t<-table(bib_df_merged$NOTES_INDEX.CA,
         bib_df_merged$NOTES_INDEX.TE)

# get agreement
t2<-round(t/sum(t),2)
ag_before <- sum(diag(t2))

# make table
knitr::kable(t,  
             caption = paste('Votes after resolving unsure discrepancies. \n
             Rows: CA votes; cols: TE votes, Agreement = ', ag_before)
             )
```

# Resolving exclude/include disagreements

## CA: exclude, TE: include

```{r}
# Resolving conflicting exclude/include annotations, part 1
IND <- which(bib_df_merged$NOTES_INDEX.CA=='exclude' & 
  bib_df_merged$NOTES_INDEX.TE=='include')
# check discrepant entries
resolved_exclude_index <- bib_df_merged$BIBTEXKEY[IND]
resolved_exclude_index # 8

# update TE votes to exclude
bib_df_merged$NOTES_INDEX.TE[bib_df_merged$BIBTEXKEY == 'aljanaki2017de'] <- 'exclude'
bib_df_merged$NOTES_INDEX.TE[bib_df_merged$BIBTEXKEY == 'li2024im'] <- 'exclude'
bib_df_merged$NOTES_INDEX.TE[bib_df_merged$BIBTEXKEY == 'pandeya2024gl'] <- 'exclude'
bib_df_merged$NOTES_INDEX.TE[bib_df_merged$BIBTEXKEY == 'sanmillancastillo2022an'] <- 'exclude'
bib_df_merged$NOTES_INDEX.TE[bib_df_merged$BIBTEXKEY == 'wang2015mo'] <- 'exclude'
bib_df_merged$NOTES_INDEX.TE[bib_df_merged$BIBTEXKEY == 'wang2016af'] <- 'exclude'
bib_df_merged$NOTES_INDEX.TE[bib_df_merged$BIBTEXKEY == 'yang2023ex'] <- 'exclude'
# update CA votes to include
bib_df_merged$NOTES_INDEX.CA[bib_df_merged$BIBTEXKEY == 'saizclar2022pr'] <- 'include'

# resolved_exclude <-c('aljanaki2017de' = 'exclude',
#                      'li2024im' = 'exclude',
#                      'pandeya2024gl' = 'exclude',
#                      'saizclar2022pr' = 'include',
#                      'sanmillancastillo2022an' = 'exclude',
#                      'wang2015mo' = 'exclude',
#                      'wang2016af' = 'exclude',
#                      'yang2023ex' = 'exclude')

```

## Summary

### CA: exclude, TE: include

| Study                       | CA                                    | TE                                                                   | Resolution/Decision                                                                                  |
|------------------|------------------|------------------|------------------|
| **aljanaki2017de**          | Multiple teams' performance reported  | Possible to report teams as substudies                               | **Exclude**: Benchmark, does not contain original study details from reporting team                  |
| **li2024im**                | No relevant task                      | Reports classification accuracy and DEAM results                     | **Exclude**: No audio features reported                                                              |
| **pandeya2024gl**           | Insufficient detail for meta-analysis | Includes timbre and global audio features, confusion matrix included | **Exclude**: Reporting on music videos; quality issues in data set                                   |
| **saizclar2022pr**          | No modeling task                      | Modeling is based on onsets                                          | **Include**: although no cross-validation, still performed task                                      |
| **sanmillancastillo2022an** | No music                              | Task present                                                         | **Exclude**: No music                                                                                |
| **wang2015mo**              | No relevant task                      | Not sure of outcome measures                                         | **Exclude**: No translation of distances into VA accuracy                                            |
| **wang2016af**              | Chapter                               | Includes relevant task                                               | **Exclude**: Meets exclusion criteria (not an article)                                               |
| **yang2023ex**              | No relevant task                      | Final metrics missing                                                | **Exclude**: Collected VA emotion ratings, but don't use audio features to predict VA (not reported) |

## TE: exclude, CA: include

```{r}
# Resolving conflicting exclude/include annotations, part 2
IND <- which(bib_df_merged$NOTES_INDEX.TE=='exclude' & 
               bib_df_merged$NOTES_INDEX.CA=='include')

# check discrepant entries
resolved_exclude2_index <- bib_df_merged$BIBTEXKEY[IND]
resolved_exclude2_index # 4

# update CA votes

bib_df_merged$NOTES_INDEX.CA[bib_df_merged$BIBTEXKEY == 'cunningham2021su'] <- 'exclude'
bib_df_merged$NOTES_INDEX.CA[bib_df_merged$BIBTEXKEY == 'eyben2015em'] <- 'exclude'
bib_df_merged$NOTES_INDEX.CA[bib_df_merged$BIBTEXKEY == 'tian2023mua'] <- 'exclude'
bib_df_merged$NOTES_INDEX.CA[bib_df_merged$BIBTEXKEY == 'tiple2022mu'] <- 'exclude'

# resolved_exclude2 <-c('cunningham2021su' = 'exclude',
#                       'eyben2015em' = 'exclude',
#                       'tian2023mua' = 'exclude',
#                       'tiple2022mu' = 'exclude')

```

## Summary

| Study                | TE                                               | CA                                               | Resolution/Decision                                          |
|------------------|------------------|------------------|------------------|
| **cunningham2021su** | Reports on IADS (not music)                      | Relevant task                                    | **Exclude**: No music                                        |
| **eyben2015em**      | Focused on laboratory singing                    | Relevant task                                    | **Exclude**: Laboratory singing of a scale                   |
| **tiple2022mu**      | Data set just includes annotation of tonic pitch | Relevant task, but only reports overall accuracy | **Exclude**: Data set not sufficiently detailed for MER task |

## Update table after resolving disagreements

```{r}
# compare raters' decisions with confusion matrix
t<-table(bib_df_merged$NOTES_INDEX.CA,
         bib_df_merged$NOTES_INDEX.TE)

# get agreement
t2<-round(t/sum(t),2)
ag_before <- sum(diag(t2))

# make table
knitr::kable(t,  
             caption = paste('Votes after resolving include vs. exclude discrepancies. \n
             Rows: CA votes; cols: TE votes, Agreement = ', ag_before)
             )
```

## Track task types

```{r}
# add column tracking task type
bib_df_merged$PARADIGM <- 'regression'
bib_df_merged[str_detect(bib_df_merged$NOTES.CA, 'classification'),]$PARADIGM <- 'classification'

# sort by task
bib_df_merged <- bib_df_merged[order(bib_df_merged$PARADIGM),]
```

## Add fields for annotation

```{r}
# add new fields for annotating bibtex library
bib_df_merged %>% mutate(emotions = ' ',
                         emotion_locus = ' ',
                         stimulus_type = ' ',
                         stimulus_duration = ' ',
                         stimulus_duration_unit = ' ',
                         stimulus_N = ' ',
                         feature_N = ' ',
                         participant_N = ' ',
                         participant_expertise = ' ',
                         participant_origin = ' ',
                         participant_sampling = ' ',
                         participant_task = ' ',
                         feature_N = ' ',
                         feature_categories = ' ',
                         feature_source = ' ',
                         feature_reduction_method = ' ',
                         model_category = ' ',
                         model_detail = ' ',
                         model_measure = ' ',
                         model_complexity_parameters = ' ',
                         model_rate_emotion_names = ' ',
                         model_rate_emotion_values = ' ',
                         model_validation = ' ',
                         final_decision = NOTES_INDEX.CA) -> bib_df_merged
```

## Export bibtex for the annotation

```{r}
bib_df_merged$NOTES_INDEX.CA <- NULL
bib_df_merged$NOTES_INDEX.TE <- NULL
```

## Write resulting bibtex library

```{r}
bib2df::df2bib(bib_df_merged, file = 'metaMER_library_template.bib')
```

Notes:

For Feature classification, use this one:

https://ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=9229494

Insert the template the annotation fields from googledoc to the bibtex file.