### 2024-05-28 studies/extraction_details.qmd
add draft doc for extraction fields

---
title: "Extraction Details"
---

To capture relevant information from studies, we expanded BiBTeX fields for each study with additional fields. For reproducibility, these instructions provide information on the process followed for each field.

# Fields

## IDENTIFIER

Unique identifier of article. Contains last name of lead author, year of publication and first two letters of article title. Hyphenated last names collapsed.

## AUTHOR

Names of all authors. Last name precedes first name and separated by comma. For multiple authors "and" precedes each listed subsequent author. E.g., Sorussa, Kanawat and Choksuriwong, Anant and Karnjanadecha, Montri

## JOURNAL

Title of journal containing article.

## NOTE

Includes number of citing articles and open access details. E.g., Cited by: 4; All Open Access, Gold Open Access, Green Open Access

## TITLE

Title of article.

## TYPE

**TODO**: Remove? All are articles

## VOLUME

Volume number of publication.

## YEAR

Publication year.

## DOI

Digital object identifier of article.

## ABSTRACT

Complete text of article abstract.

## PUBLICATION_STAGE

**TODO:** Remove? All final.

## SOURCE

Database article was sourced from. Scopus, Web of Science (WoS) or OpenAlex.

## AUTHOR_KEYWORDS

Corresponding keywords for article indicated by author.

## PARADIGM

Paradigm used for MER task. Either classification or regression.

## NOTES_AUTHORINITIALS

Decision and comments by respective author

## EMOTIONS

List of emotion names or labels separated by commas. **TODO:** Discuss potential redundancies with `MODEL_RATE_EMOTION_NAMES`

## EMOTION_LOCUS

Perceived or felt **TODO:** Remove? All appear to be perceived

## STIMULUS_TYPE

Metadata pertaining to stimuli employed in paradigm. Can be listed as genres of music stimuli employed, or if stimuli come from a standard database, name of standard.

## STIMULUS_DURATION

Duration of stimuli, if applicable. Unit of measurement (seconds, measures) specified in `STIMULUS_DURATION_UNIT`

## STIMULUS_DURATION_UNIT

Unit of measurement pertaining to `STIMULUS_DURATION`. E.g., seconds, measures, etc.

## STIMULUS_N

Number of stimuli employed in experiment.\
**TODO** Discuss handling of subset selection, train, test, etc

## FEATURE_N

Number of features included in model. **TODO** Discuss handling of subset selection, train, test, etc

## PARTICIPANT_N

Total number of participants in experiment. **TODO** Discuss handling of secondary information. E.g., MediaEval

## PARTICIPANT_EXPERTISE

Expertise of annotators. E.g., experts, non-experts, not specified.

## PARTICIPANT_ORIGIN

Origin country of participants, or online platform participants were recruited from (e.g., MTurk)

## PARTICIPANT_SAMPLING

How participants were recruited (e.g., convenience, random sampling, crowdsourcing)

## PARTICIPANT_TASK

Nature of rating/classification task undertaken by participants. E.g., rate, annotate. **TODO** discuss multi-stage studies. E.g., MediaEval categories validated by participants using AV ratings.

## FEATURE_CATEGORIES

Names of categories analyzed features pertain to, based on names in Panda (2021). Includes names of all pertinent categories: *Melody*, *Rhythm*, *Timbre*, *Pitch*, *Tonality*, *Expressivity*, *Texture*, *Form*, *Vocal*, *High-Level* **TODO** Discuss feasibility of this approach. Other nominalizations/categories not covered (e.g., spectral, temporal, lyric)

## FEATURE_SOURCE

Name(s) of feature analysis toolboxes. **TODO** Discuss original features (as in `Panda2020no`)

## FEATURE_REDUCTION_METHOD

Name(s) of feature reduction or feature selection methods employed.

## MODEL_CATEGORY

Name of model type. One of regression, classification *TODO* Discuss redundancy with `PARADIGM`. Consider cases where regression and classification employed; or regression approach to classification (e.g., `sorussa2020em`)

## MODEL_DETAIL

Additional information pertaining to predictive model, such as the name of algorithm used and other pertinent parameters. E.g., RandomForest, Commonality Analysis, Multiple Regression, Neural Networks, LDSM, etc. *TODO* Discuss how much information to include.

## MODEL_MEASURE

Metric used in model evaluation. E.g., 'R2' for $R^2$, MSE, CCC, Classification accuracy, etc.

## MODEL_COMPLEXITY_PARAMETERS

Additional information pertaining to predictive model. Format as R code when encoding multiple pertinent parameters. E.g., `c(training_epochs = 100, layer_n = c(1,2), lstm_units = c(124, 248))`

## MODEL_RATE_EMOTION_NAMES

Names of predicted emotions. E.g., valence, arousal, happy, sad, angry, fearful, etc.

## MODEL_RATE_EMOTION_VALUES

Pertinent prediction of model summaries. Report as *R* arrays, including summary statistics in variables. When reporting results of multiple models, concatenate multiple entries with `rbind`. When reporting results for different toolboxes or feature subsets, assign each to a new BiBTeX field with relevant identifier following final underscore.

\### EXAMPLES:

#### Regression

Report relevant summary statistics in R code. Different feature distinguished through BiBTeX fields (`MODEL_RATE_EMOTION_VALUES_MFCC`, `MODEL_RATE_EMOTION_VALUES_ESSENTIA` ). Fitted models concatenated with `rbind` (`linear.regression`, `SMOreg`, `RNN1.124LSTM`, etc.). Model results reported with names in first array (`R2_arousal`, `MAE_arosal`, `R2_valence`, `MAE_valence`), and recycled in each subsequent model.

```         
MODEL_RATE_EMOTION_VALUES_MFCC = {
  rbind(
    linear.regression = c(R2_arousal = 0.61, MAE_arousal = 0.13, R2_valence = 0.07, MAE_valence = 0.17),
    SMOreg = c(0.6,0.13,0.10,0.17),
    RNN1.124LSTM = c(0.58,0.14,0.14,0.17),
    RNN2.124.124LSTM = c(0.66,0.12,0.11,0.16),
    RNN3.248LSTM = c(0.61,0.13,0.13,0.16),
    RNN4.248.248LSTM = c(0.64,0.12,0.14,0.15)
  )
},
MODEL_RATE_EMOTION_VALUES_ESSENTIA = {
  rbind(
    linear.regression = c(R2_arousal = 0.07, MAE_arousal = 0.25, R2_valence = 0.06, MAE_valence = 0.19),
    SMOreg = c(0.48,0.18,0.27,0.17),
    RNN1.124LSTM = c(0.54,0.14,0.21,0.16),
    RNN2.124.124LSTM = c(0.67,0.12,0.32,0.13),
    RNN3.248LSTM = c(0.58,0.14,0.32,0.14),
    RNN4.248.248LSTM = c(0.69,0.11,0.40,0.13),
    RNN5.529LSTM = c(0.61,0.13,0.29,0.15),
    RNN6.529.529LSTM = c(0.68,0.12,0.36,0.14)
  )
}
```

#### Classification

Report confusion matrices as *R* arrays. The following example classifies emotions into five categories for multiple feature subsets. Different feature distinguished through BiBTeX fields (`MODEL_RATE_EMOTION_VALUES_UNAMBIGUOUS_SMO`, `MODEL_RATE_EMOTION_VALUES_FULL_SMO`, etc.). Confusion matrix classes concatenated with `rbind` (`C1`, `C2`, `C3`, etc.). Classification results reported with names in first array (`C1`, `C2`, `C3`, etc.), and recycled in each subsequent model.

```         
MODEL_RATE_EMOTION_VALUES_UNAMBIGUOUS_SMO = {
rbind(
    C1 = c('C1' = 52.6, 'C2' = 17.1, 'C3' = 0.0, 'C4' = 9.2, 'C5' = 21.1),
    C2 = c(12, 65.2, 7.6, 4.3, 10.9),
    C3 = c(1.1,9.9,73.6,14.3,1.1),
    C4 = c(3.6,10.8,16.9,61.5,7.2),
    C5 = c(21.9,19.2,0,6.8,52.1)
 )
}
MODEL_RATE_EMOTION_VALUES_FULL_SMO = {
rbind(
    C1 = c('C1' = 56.0, 'C2' = 19, 'C3' = 0.0, 'C4' = 7, 'C5' = 18),
    C2 = c(11.3,58.9,13.7,4,12.1),
    C3 = c(0,13.2,68.6,18.2,0),
    C4 = c(4.9,11.7,22.5,54.2,6.7),
    C5 = c(18.5,12.2,1,12.2,56.1)
  )
},
MODEL_RATE_EMOTION_VALUES_FULL_FULLPOLYGONAL = { rbind(
  C1 = c('C1' = 65, 'C2' = 11, 'C3' = 12, 'C4' = 0, 'C5' = 12),
  C2 = c(14.5, 62.1, 22.6, 0.8, 0),
  C3 = c(0, 13.2, 68.6, 18.2, 0),
  C4 = c(0.8, 0, 20.8, 54.2, 24.2),
  C5 = c(38.8, 1, 0, 8.2, 52)
)
},
MODEL_RATE_EMOTION_VALUES_FULL_REDUCEDPOLYGONAL = { 
rbind(
  C1 = c('C1' = 67, 'C2' = 16, 'C3' = 4, 'C4' = 0, 'C5' = 13),
  C2 = c(21.8,55.6,22.6,0,0),
  C3 = c(0,9.1,77.7,13.2,0),
  C4 = c(0,0,23.3,55,21.6),
  C5 = c(27.5,0,0,10.2,62.3)
)
}
```

## MODEL_VALIDATION

Validation method used (if applicable). E.g., 10-fold cross validation. Leave one out cross validation. **TODO**: Discuss train/test/validation formatting. When using external validation set, report name? E.g., AcousticBrainz (`alvarez2023ri`)