### 2024-06-20 index.qmd

# README

How well we can predict emotions in music? What is the evidence in the published literature for explaining what emotions the listeners can perceive in music when the source consists of audio examples. To what degree the results are dependent on the actual models, emotions, musical/acoustic features, or musical materials or participants?

To obtain answers to these questions, we have set out to record and analyse the current state of the art from the literature using a meta-analysis paradigm. We focus on Music Emotion Recognition and hence the acronym `metaMER`.

The public-facing version of the repository is available at
[https://tuomaseerola.github.io/metaMER/](https://tuomaseerola.github.io/metaMER/)

## Plan

We define the aims and methods in [preregistration plan](preregistration/preregistration.qmd), which has beeb preregistered at [https://osf.io/6y3dr/](https://osf.io/6y3dr/).

### Study Search and Selection

Search databases and criteria are documented in [studies/search_syntax.qmd](studies/search_syntax.qmd).

### Data Extraction and Coding

Data extraction is described in [extraction details](studies/extraction_details.qmd). See also [pass3 comparison ](studies/pass3_comparison.qmd).

The data will be parsed to a tabular format using a custom [library_parser.qmd](studies/library_parser.qmd).

### Analysis

Data analysis is covered in [analysis/analysis.qmd](analysis/analysis.qmd) document.

### Manuscript

The study report is available [manuscript/manuscript.qmd](manuscript/manuscript.qmd) document.