### 2024-09-25 manuscript/_figure1.qmd

```{r}
#| label: fig1
#| fig.height: 7
#| echo: false
#| fig.cap: > 
#|   Forest plot of valence predictions from the MER models.

library(metafor)
library(dmetar)
library(meta)

R_summary <- read.csv("../analysis/R_summary.csv")
tmp <- dplyr::filter(R_summary,dimension=="valence")
tmp$studyREF[tmp$studyREF=="Wang et al 2022"] <- c("Wang et al. 2022a","Wang et al. 2022b")

# Max values
m.cor <- metacor(cor = valuesMax,     # values 
                 n = stimulus_n,
                 studlab = studyREF,#citekey, # unique_id
                 data = tmp,
                 fixed = FALSE,
                 random = TRUE,
                 prediction = TRUE,
                 #                 backtransf = TRUE,
                 #                 sm = "ZCOR",
                 method.tau = "REML",# could be PM (Paule-Mandel) as well
                 method.random.ci = "HK", 
                 title = "MER: Regression: Valence: Summary")

forest(m.cor,
       sortvar = TE,
       fontfamily = "Times",
       hetstat =FALSE,
       prediction = FALSE, 
       print.tau2 = FALSE,
       leftlabs = c("Author", "g", "SE"),layout = "meta",addrows =0,rows.gr=0,col.diamond='black') # RevMan5

```