### 2024-10-17 manuscript/_figure1.qmd
new figures in a separate quarto docs

```{dot}
//| label: fig1
//| fig-cap: "Flowchart of the study inclusions/eliminations."

digraph sub {
  node [fontname = Helvetica, fontsize = 14, shape = box, width = 4]
  A [label = "Records identified from:
                Web of Science (n = 142)
                Scopus (n = 227)
                OpenALEX (n = 278)" ]
  B [label = "Records removed before screening: 
                Duplicates entries/errors (n = 94)" ]
  C [label = "Records after duplicates removed (n = 553)"]
  D [label = "Records screened (n = 553)"]
  E [label = "Records removed: 
                Irrelevant titles/themes (n = 338)
                Abstract screening (n = 95)
                Discussion (n = 24)"]
  F [label = "Full text articles assessed (n = 96)"]
  G [label = "Studies included in final review (n = 46)"]
  H [label = "Full text articles excluded
             (n = 50)"]
  I [label = "Studies excluded in final review (n = 10)"]
  J [label = "Final selection (n = 36)"]
  A->B
  A->C
  C->D
  D->E
  D->F
  F->H 
  F->G
  G->I
  G->J

  { rank = same; A B}
  { rank = same; D, E}
  { rank = same; F, H}
  { rank = same; G, I}
}
```