### 2024-10-17 manuscript/_figure2.qmd

```{r}
#| label: fig2
#| fig-width: 16
#| fig-height: 6
#| echo: false
#| fig.cap: > 
#|   Forest and funnel plots of valence predictions from the MER models.

library(metafor, warn.conflicts = FALSE)
library(dmetar,warn.conflicts = FALSE)
library(meta,warn.conflicts = FALSE)
library(DescTools,warn.conflicts = FALSE)
library(ggplot2,quietly = TRUE)
library(ggrepel,quietly = TRUE)

R_summary <- read.csv("../analysis/R_summary.csv")
tmp <- dplyr::filter(R_summary,dimension=="valence")
tmp$studyREF[tmp$studyREF=="Wang et al 2022"] <- c("Wang et al. 2022a","Wang et al. 2022b")

# Max values
m.cor <- metacor(cor = valuesMax,     # values 
                 n = stimulus_n,
                 studlab = studyREF,#citekey, # unique_id
                 data = tmp,
                 fixed = FALSE,
                 random = TRUE,
                 prediction = TRUE,
                 #                 backtransf = TRUE,
                 #                 sm = "ZCOR",
                 method.tau = "REML",# could be PM (Paule-Mandel) as well
                 method.random.ci = "HK", 
                 title = "MER: Regression: Valence: Summary")

#### Forest plot using forestplot
library(forestplot)
data<-tibble::tibble(mean=m.cor$cor,lower=DescTools::FisherZInv(m.cor$lower),upper=DescTools::FisherZInv(m.cor$upper),study=m.cor$studlab,n=m.cor$n,cor=round(m.cor$cor,2))
data<-dplyr::arrange(data,mean)

fp1 <- grid.grabExpr(print(data |>
  forestplot(labeltext = c(study, n, cor),
             xlab = "Correlation",
             xticks = c(0, .25,.5,.75, 1),
             clip = c(0, 1))|>
    fp_add_header(study = "Study",n = "N",cor = expression(italic(r))) |>
    fp_append_row(mean  = m.cor$TE.common,
                lower = m.cor$lower.common,
                upper = m.cor$upper.common,
                study = "Summary",
                n = sum(m.cor$n),
                cor = round(m.cor$TE.common,2),
                is.summary = TRUE) |>
  fp_set_style(box = "grey50",
               line = "grey20",
               summary = "black",
                txt_gp = fpTxtGp(label = list(gpar(cex = 0.80)),
                                ticks = gpar(cex = 0.80),
                                xlab  = gpar(cex = 0.80)))|>
    fp_decorate_graph(grid = structure( m.cor$TE.common,gp = gpar(lty = 2, col = "grey30")))
)
)

#### Funnel plot using a custom function
source('../etc/custom_funnel_plot.R')
fp2 <- suppressWarnings(custom_funnel_plot(m.cor))

gridExtra::grid.arrange(fp1, fp2, ncol=2, widths=c(2,1))

```