### 2025-02-18 manuscript/_datasets.qmd

```{r}
#| label: datasets
```

<!-- This is a reduced table from /etc/Secondary Databases.md -->

| Dataset                                                                                                | Stim. Type                 | Stim. Dur. (s)  | Stim. N          | Feature N    | Ppt. N          | Feature Source                                        | In studies                                                                                                               |
| ------------------------------------------------------------------------------------------------------ | -------------------------- | --------------- | ---------------- | ------------ | --------------- | ----------------------------------------------------- | ------------------------------------------------------------------------------------------------------------------------ |
| [**MediaEval**](https://cvml.unige.ch/databases/emoMusic/)                                             | Western pop                | 45              | 744              | 6669         | 10/track        | OpenSMILE                                             | @bai2016di, @bai2017mu, @yang2021an, @chin2018pr, @coutinho2017sh, @markov2014mu, @medina2020em, @wang2022co, @xie2020mu |
| [**DEAM**](https://cvml.unige.ch/databases/DEAM/)                                                      | Pop                        | 45              | 1802             | 260          | 5-10/track      | OpenSMILE                                             | @sorussa2020em, @orjesek2022en, @panwar2019ar, @zhang2023mo                                                              |
| [**AMG1608**](https://github.com/loichan-tw/AMG1608_release)                                           | Pop                        | 30              | 1608             | 72           | 643             | MIR Toolbox, YAAFE                                     | @chen2017co, @hu2017cr, @wang2022cr                                                                                      |
| [**EMOPIA**](https://annahung31.github.io/EMOPIA/)                                                     | Piano Solo (pop music)     | 30-40           | 387              | 24           | 1 annot./track  | MIDI Toolbox                                          | @bhuvanakumar2023em                                                                                                      |
| [**NTUMIR**](https://web.archive.org/web/20170510081611/mac.citi.sinica.edu.tw/~yang/MER/NTUMIR-60)    | Famous pop songs           | 25              | 60               | 46           | 40 annot./track | MIR Toolbox, Sound Description Toolbox, MA Toolbox     | @chin2018pr                                                                                                              |
| [**Soundtracks**](https://osf.io/p6vkg/wiki/home/)                                                     | Obscure film soundtracks   | 15              | 110              | NA           | 116             | NA                                                    | @wang2022co                                                                                                              |
| [**PSIC3839**](https://github.com/xl2218066/PSIC3839)                                                  | Chinese popular            | 180             | 3839             | NA           | 87              | Librosa                                               | @xu2021us                                                                                                                |
| [**CH818**](ccmir.cite.hku.hk/data/)                                                                   | Chinese pop                | 30              | 818              | 15           | 3               | MIR Toolbox, PsySound, Chroma Toolbox, Tempogram Toolbox| @hu2017cr                                                                                                                |
| [**Zhang et al. (2015)**](LINK)                                                                        | Chinese pop                | 30              | 171              | 84           | 10              | MA Toolbox, MIR Toolbox, Coversongs                     | @zhang2016br                                                                                                             |
| [**PMEmo**](http://huisblog.cn/PMEmo/)                                                                 | Pop songs                  | Variable        | 794              | 6373         | 457             | ComParE 2013 baseline feature set                     | @zhang2023mo                                                                                                             |
| [**NJU-V1**](https://cs.nju.edu.cn/sufeng/data/musicmood.htm)                                          | Limited detail             | Variable        | 777              | Not reported | NA (tags)       | NA                                                    | @agarwal2021an                                                                                                           |
| [**ISMIR-2012**](http://yadingsong.blogspot.com/2015/03/popular-music-emotion-dataset-ismir2012.html)  | Popular music              | 30 or 60        | 2904             | 54           | NA (tags)       | MIR Toolbox                                            | @agarwal2021an                                                                                                           |
| [**MIREX2009**](https://www.music-ir.org/mirex/wiki/2009:Music_Structure_Segmentation_Results)         | Popular                    | Full            | 297              | 3            | NA              | Paulus & Klapuri (2009)                               | @yeh2014po                                                                                                               |
| [**Million Songs Dataset**](http://millionsongdataset.com/pages/getting-dataset/)                      | Pop                        | Full            | 1,000,000        | 55           | None            | EchoNest                                              | @cao2023th                                                                                                               |
| [**Free Music Archive**](https://freemusicarchive.org/home)                                            | Various                    | Variable        | >100,000         | NA           | NA              | NA                                                    | @koh2023me                                                                                                               |
| [**Jamendo**](https://www.music-ir.org/mirex/wiki/2015:Main_Page)                                      | Various                    | Variable        | 10,000           | 24           | NA              | Metadata                                              | @hu2022de                                                                                                                |
| **Chinese Classical Music Dataset**                                                                    | Chinese classical          | ~30s            | 500              | 557          | 20              | Essentia, MIR Toolbox                                  | @wang2022co                                                                                                              |
Notes: $\dagger$ Used in @alvarez2023ri

<!-- I TOOK THESE OUT FOR NOW
| [**Spotify API**](https://developer.spotify.com/documentation/web-api)                                 | Various                    | Variable        | 5192$\dagger$    | 12           | NA              | EchoNest                                              | @alvarez2023ri                                                                                                           |
| [**Acoustic Brainz**](https://acousticbrainz.org/)                                                     | Various                    | Variable        | 60,000$\dagger$  | Not reported | NA (tags)       | Essentia                                              | @alvarez2023ri                                                                                                           |
-->