### 2025-11-28 analysis/supp_tables.qmd
Updated manuscript and bibtex file (R1)

# Supporting tables

Updated 28/11/2025

```{r}
#| warning: false
#| eval: true
#| echo: false
#library(dmetar,quietly = TRUE)
library(tidyverse,quietly = TRUE)
#library(meta)
#library(DescTools)
#library(ggrepel)
#library(forestplot)
```

#### Summarise all models

```{r}
#| warning: false
#| eval: true
#| id: table-summary1
#| output: asis

R_studies <- read.csv("R_studies.csv")
R_studies$unique_id_task <- paste(R_studies$unique_id,R_studies$dimension,sep="-")
#length(unique(R_studies$unique_id_task))
RS <- reframe(group_by(R_studies,model_class_id,),id=unique_id_task)
colnames(RS) <- c("Technique","Study-Model ID")

C_studies <- read.csv("C_studies.csv")
CS <- reframe(group_by(C_studies,model_class_id,),id=unique_id)
colnames(CS) <- c("Technique","Study-Model ID")

print(knitr::kable(RS,caption="Regression studies by model class"))
print(knitr::kable(CS,caption="Classification studies by model class"))
```